---
title: "Launching and Using the Refuginator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Launching and Using the Refuginator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The Refuginator is a dashboard designed to facilitate the identification of historical and geohistorical refugia using the methods proposed by Morley et al. (2025). This vignette provides users a comprehensive walkthrough on how to launch and use the Refuginator dashboard using the `refuginator` package in R.

## Sections

*[Getting Started]


## Getting Started

Users can install the `refuginator` package from CRAN with the command:

```{r, eval = FALSE}
install.packages("refuginator")
```

Once the package is installed, users can then launch the Refuginator dashboard by attaching the `refuginator` library and using the `launchRefuginator()` command:

```{r, eval = FALSE}
library(refuginator)
launchRefuginator()
```

Or, if users prefer to attach the package directly in the function's namespace:

```{r, eval = FALSE}
refuginator::launchRefuginator()
```


## 
